---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

This project has not need to load any package.

## Loading and preprocessing the data
Make sure the data file had been extracted to the working directory

```{r, echo=TRUE}
act <- read.csv("activity.csv")
```
## What is mean total number of steps taken per day?
Draw the histogram of the total number of steps taken per day:
Missing values (NA) are excluded in this step.
```{r}
hist(
  tapply(act$steps, act$date, sum, na.rm=T)
     , main = "total number of steps taken per day")
```

Here are the mean and median total number of steps taken per day
```{r}
summary(tapply(act$steps, act$date, sum, na.rm=T))[c(3, 4)]
```

## What is the average daily activity pattern?

A time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).
```{r}
plot(tapply(act$steps, act$interval, mean, na.rm = T) ~ levels( as.factor(act$interval) ), type = "l",
     ylab = "average steps", xlab = "5-min interval")
abline(h = mean(tapply(act$steps, act$interval, mean, na.rm = T)), col="red")
```

The 5-min interval having the maximum number of steps is...
```{r}
rownames( sort(tapply(act$steps, act$interval, mean, na.rm = T), decreasing = T) )[1]
```

## Imputing missing values

The total number of missing values is ...
```{r}
sum(is.na(act$steps) == TRUE)
```

Filling in all of the missing values with the average number of steps
```{r}
tmp = act$steps
tmp[is.na(act$steps) == TRUE] = rep(tapply(act$steps, act$interval, mean, na.rm = T), length(act$date[is.na(act$steps) == TRUE])/length(tapply(act$steps, act$interval, mean, na.rm = T)))
```

Create a new dataset that the missing values are replaced.
```{r}
act_n <- data.frame(steps = tmp, date = act$date, interval = act$interval)
```

Calculate total number of steps taken per day. 
Missing values are replaced by  the average number of steps.
```{r}
steps_n_total <- tapply(act_n$steps, act_n$date, sum, na.rm=T)
```

Draw the histogram after missing data replaced:
```{r}
hist(tapply(act_n$steps, act_n$date, sum, na.rm=T), 
     main = "total number of steps taken per day")
```

The mean and median total number of steps taken per day after missing data replaced.
```{r}
summary(tapply(act_n$steps, act_n$date, sum, na.rm=T))[c(3, 4)]
```

The distribution (mean and median) differ from original data
```{r}
summary(tapply(act_n$steps, act_n$date, sum, na.rm=T))[c(3, 4)] == 
  summary(tapply(act$steps, act$date, sum, na.rm=T))[c(3, 4)]
```

Before replaced missing data, the skewness of distribution is positive:
```{r}
summary(tapply(act$steps, act$date, sum, na.rm=T))[3] - 
  summary(tapply(act$steps, act$date, sum, na.rm=T))[4]
```
After replaced missing data, the histogram is like normal distribution
```{r}
summary(tapply(act_n$steps, act_n$date, sum, na.rm=T))[3] - 
  summary(tapply(act_n$steps, act_n$date, sum, na.rm=T))[4]
```


## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable **WEEK** in the dataset with two levels: "weekday" and "weekend" indicate whether a given date is a weekday or weekend day.
**Note: this markdown is completed in Taiwan, timezone = CST**
```{r}
WEEK <- ifelse(
   ((weekdays( as.Date(act_n$date) ) == "星期六") | (weekdays( as.Date(act_n$date) ) == "星期日") ) == TRUE,
   "weekend",
   "weekday"
)
```

Make a panel plot containing the time series plots across all weekend days and all weekday days.
```{r, fig.height=10}
par(mfrow=c(2, 1))

plot(tapply(act_n$steps[WEEK=="weekend"], act_n$interval[WEEK=="weekend"], mean, na.rm = T) ~ levels( as.factor(act_n$interval[WEEK=="weekend"]) ), type = "l",
     ylab = "average steps", xlab = "", main = "weekend")

plot(tapply(act_n$steps[WEEK=="weekday"], act_n$interval[WEEK=="weekday"], mean, na.rm = T) ~ levels( as.factor(act_n$interval[WEEK=="weekday"]) ), type = "l",
     ylab = "average steps", xlab = "5-min interval", main = "weekday")
```